Network Configuration Synthesis with Abstract Topologies

### 2016-11-07 paper/abstraction/paper.tex
small wording fixes to the first three sections

\textbf{Abstract---}
We develop techniques for automated reasoning about end-to-end properties of abstract network topologies. We extend a network synthesis tool, \sysname, to allow network operators to write a high-level routing policy over an abstraction of their network topology. From such a specification, \sysname synthesizes a collection router configurations running the distributed Border Gateway Protocol (BGP) that are guaranteed to implement the centralized routing policy for any concrete instantiation of the abstract topology, and for any combination of network failures in the resulting concrete topology.
%
Additionally, properties proven by \sysname over the abstract topology, such as reachability between a source and destination under k-failures, are also guaranteed to hold under any concrete instantiation of the abstract topology.
%
Something something, expandable/incremental networks. Something something scales compilation to large networks. Something something, templates easier to understand.
%These strong guarantees allow operators to reason about the routing behavior of their network at a high-level abstraction, while also generating an efficient low-level implementation.
%
%In addition, operators are able to safely change or expand their network in any way, so long as the new resulting topology also respects the abstraction.
%
We test \sysname out on ... blah blah blah

@@

Unfortunately, such templates have no particular semantics other than the
semantics
that one obtains from instantiating a template's parameters.
Moreover,
template construction is a low-level, manual process:  
network engineers do their
best to design templates using assembly-language-like primitives.
They hope their templates
that will work properly, test them using simulations, but
there is no support for verification of key network-wide properties 
such as connectivity and fault tolerance. In addition,
network engineers have no tools to help them design for network
evolution.  For instance, the process of adding new equipment to
expand the capacity of data center may not be compatible with the original
templates, or may cause a trickle-down of changes to be made network-wide.
Each change to a router configuration typically incurs
some risk of temporary instable in a network, so minimizing such changes
is an important design goal.  
%It is no wonder that a major
%network outage, such as the one that grounded all United Air flights
%in July 2015~\cite{x}, makes international news every few months.

To combat these problems, we present \sysname, a
new high-level language designed to
provide improved support for the configuration of large, structured networks.
\sysname allows network operators to define \emph{abstract network topologies}
that
classify devices in terms of their roles and how they are connected
to one another.  For instance, a programmer may
define separate roles for ``top-of-rack router'' and
``tier-one aggregation router'' and specify that every top-of-rack router
is connected to at least two different tier-one aggregation routers,
thereby providing a degree of fault tolerance of one aggregation router
were to fail.

@@

We demonstrate the difficulty of configuring networks using the example data center network in Figure~\ref{fig:example}. The boxes denote routers. Using terminology for fat tree networks~\cite{fattree}, T[1--8] denote top-of-rack (ToR) routers, A[1--8] denote aggregation routers, and S[1--2] denote spine routers. The spine routers are connected to the
rest of the Internet through Peer[1--2]. The aggregation routers
bridge the spine to the ToRs routers. The ToR routers attach to a set of servers (``a rack'') that host services with address prefixes P[1--8].

@@

To simplify network configuration, the operations community is adopting a template-based approach~\cite{hatch,thwack}. Instead of authoring a configuration file per router, operators author one configuration template per role.
%
A {\em role} refers to specific functionality and is played by one or more routers.\footnote{If a role is played by devices from multiple vendors, there is one template per role-vendor combination because templates use vendor configuration language.}
%
The network in Figure~\ref{fig:example} has five roles: spine, global aggregator, global ToR, local aggregator, and local ToR.
%
Templates have parameters for various aspects of the configuration (\EG, interface addresses, neighbor list) and are compiled to device configurations by instantiating the parameters using a database of network information.

@@

Consider the network in Figure~\ref{fig:example2}, which is similar to Figure~\ref{fig:example}; it has the same five roles, connected in a similar hierarchy. One might think that the same templates (with different database entries) can be used for both topologies, but that is not true. For instance, consider templates that disallow ``valley" paths, which go up, down, and then up again; it is a common to disallow valley paths in a data center because they are longer than shortest paths. These templates will work perfectly for Figure~\ref{fig:example}, but violate the fault tolerance requirement ($p7$) when used for Figure~\ref{fig:example2}.  Specifically, in Figure~\ref{fig:example2}, an aggregation-induced blackhole (\S\ref{sec:motivation}) will occur when the link S1--A1 fails; after this failure, S1 has no valley-free path to P[1--2] even though it will continue to get traffic for these prefixes as it announces  the covering prefix PG (because it gets routes for P[3--4]). In contrast, in Figure~\ref{fig:example}, no such blackhole will occur after a single link failure because spine routers have two links to each pod.

@@

\para{Configuration synthesis with concrete topologies}
Many recent systems generate router configurations using high-level policy statements~\cite{narain:lisa05,narain+:configassure,dadc,propane} and the network's concrete topology.
%These systems eliminate entire classes of configuration errors by construction, and, in the case of
%Propane~\cite{propane}, guarantee that high-level policy is implemented correctly, even in the presence of
%arbitrary combinations of failures.
They are able to provide strong guarantees on configuration correctness but 
our conversations with operators at two major cloud providers reveal that they have two limitations that discourage adoption.  The first is abstraction mismatch. Operators of large networks think of the network, and to develop management tools, using \emph{roles}, and not individual devices. They are naturally reluctant to use a system that does not
understand roles as a first-order abstraction. If they want to debug, analyze or
understand system output, they will have to consider
\emph{hundreds} of independent device configurations instead of just
\emph{a handful} of role configurations. Unfortunately, even if two devices play the same role,
there is no guarantee a concrete synthesis system will generate (syntactically) similar configurations.

>>>>

\textbf{Abstract---}
We develop techniques for automated reasoning about end-to-end properties of abstract network topologies. We develop a network synthesis tool, \sysname, to allow network operators to write a high-level routing policy over an abstraction of their network topology. From such a specification, \sysname synthesizes a collection of router configurations running the distributed Border Gateway Protocol (BGP) that are guaranteed to implement the centralized routing policy for any concrete instantiation of the abstract topology, and for any combination of network failures in the resulting concrete topology.
%
Additionally, properties proven by \sysname over the abstract topology, such as reachability between a source and destination under k-failures, are also guaranteed to hold under any concrete instantiation of the abstract topology.
%
Something something, expandable/incremental networks. Something something scales compilation to large networks. Something something, templates easier to understand.
%These strong guarantees allow operators to reason about the routing behavior of their network at a high-level abstraction, while also generating an efficient low-level implementation.
%
%In addition, operators are able to safely change or expand their network in any way, so long as the new resulting topology also respects the abstraction.
%
We test \sysname out on ... blah blah blah

@@

Unfortunately, such templates have no particular semantics other than the
semantics
that one obtains from instantiating a template's parameters.
Moreover,
template construction is a low-level, manual process:  
network engineers do their
best to design templates despite the assembly-language-like primitives for network configuration.
They hope their templates
will work properly and test concrete instantiations of them using simulations, but
there is no support for verification of key network-wide properties 
such as connectivity and fault tolerance. In addition,
network engineers have no tools to help them design for network
evolution.  For instance, the process of adding new equipment to
expand the capacity of data center may not be compatible with the original
templates, or may cause a trickle-down of changes to be made network-wide.
Each change to a router configuration typically incurs
some risk of temporary instability in a network, so minimizing such changes
is an important design goal.  
%It is no wonder that a major
%network outage, such as the one that grounded all United Air flights
%in July 2015~\cite{x}, makes international news every few months.

To combat these problems, we present \sysname, a
new high-level language designed to
provide improved support for the configuration of large, structured networks.
\sysname allows network operators to define \emph{abstract network topologies}
that
classify devices in terms of their roles and how they are connected
to one another.  For instance, a programmer may
define separate roles for ``top-of-rack router'' and
``tier-one aggregation router'' and specify that every top-of-rack router
is connected to at least two different tier-one aggregation routers,
thereby providing a degree of fault tolerance if one aggregation router
were to fail.

@@

We demonstrate the difficulty of configuring networks using the example data center network in Figure~\ref{fig:example}. The boxes denote routers. Using terminology for fat tree networks~\cite{fattree}, S[1--2] denote spine routers, A[1--8] denote aggregation routers, and T[1--8] denote top-of-rack (ToR) routers. The spine routers are connected to the
rest of the Internet through Peer[1--2]. The aggregation routers
bridge the spine to the ToR routers. The ToR routers attach to a set of servers (``a rack'') that host services with address prefixes P[1--8].

@@

To simplify network configuration, the operations community is adopting a template-based approach~\cite{hatch,thwack}. Instead of authoring a configuration file per router, operators author one configuration template per role.
%
A {\em role} refers to specific functionality and is played by one or more routers.\footnote{If a role is played by devices from multiple vendors, there is one template per role-vendor combination because templates use vendor-specific configuration languages.}
%
The network in Figure~\ref{fig:example} has five roles: spine, global aggregator, global ToR, local aggregator, and local ToR.
%
Templates have parameters for various aspects of the configuration (\EG, interface addresses, neighbor list) and are compiled to device configurations by instantiating the parameters using a database of network information.

@@

Consider the network in Figure~\ref{fig:example2}, which is similar to Figure~\ref{fig:example}; it has the same five roles, connected in a similar hierarchy. One might think that the same templates (with different database entries) can be used for both topologies, but that is not true. For instance, consider templates that disallow ``valley" paths, which go up, down, and then up again; it is a common to disallow valley paths in a data center because they are longer than shortest paths. These templates will work perfectly for Figure~\ref{fig:example} but silently violate the fault tolerance requirement ($p7$) when used for Figure~\ref{fig:example2}.  Specifically, in Figure~\ref{fig:example2}, an aggregation-induced blackhole (\S\ref{sec:motivation}) will occur when the link S1--A1 fails; after this failure, S1 has no valley-free path to P[1--2] even though it will continue to get traffic for these prefixes as it announces  the covering prefix PG (because it gets routes for P[3--4]). In contrast, in Figure~\ref{fig:example}, no such blackhole will occur after a single link failure because spine routers have two links to each pod.

@@

\para{Configuration synthesis with concrete topologies}
Many recent systems generate router configurations using high-level policy statements~\cite{narain:lisa05,narain+:configassure,dadc,propane} and the network's concrete topology.
%These systems eliminate entire classes of configuration errors by construction, and, in the case of
%Propane~\cite{propane}, guarantee that high-level policy is implemented correctly, even in the presence of
%arbitrary combinations of failures.
They are able to provide strong guarantees on configuration correctness but 
our conversations with operators at two major cloud providers reveal that they have two limitations that discourage adoption.  The first is abstraction mismatch. Operators of large networks think of the network, and develop management tools, using \emph{roles}, and not individual devices. They are naturally reluctant to use a system that does not
understand roles as a first-order abstraction. If they want to debug, analyze or
understand system output, they will have to consider
\emph{hundreds} of independent device configurations instead of just
\emph{a handful} of role configurations. Unfortunately, even if two devices play the same role,
there is no guarantee a concrete synthesis system will generate (syntactically) similar configurations.
